---
title: "Etherum"
author: "Name"
date: "`r Sys.Date()`"
output: pdf_document
---

# Packages
```{r}
library(tidyverse)
library(fpp2)
```


# Importing Data
```{r}
etherum <- read_csv("etherum.csv") |> 
  rename(ClosingPrice = `closing price`)
head(etherum)
```


# Data Cleaning
```{r}
transformed_etherum <- etherum <- etherum |> 
  mutate(
    MonthName = lubridate::month(date,label = TRUE),
         Year = lubridate::year(date)
  )

etherum$MonthName <- fct_relevel(etherum$MonthName,levels(etherum$MonthName))
```


# Exploratory Data Analysis

### Average price of Etherum for each year under each month
```{r}
etherum |> 
  group_by(Year, MonthName) |> 
  summarize(Avg_ClosingPrice = mean(ClosingPrice)) |>
  pivot_wider(names_from = Year, values_from = Avg_ClosingPrice) |> 
  arrange(MonthName) |> 
  write_csv("etherum_w.csv")
```


### Barplot showing Average price of Etherum In each year
```{r}
etherum |> 
  group_by(Year, MonthName) |> 
  summarize(Avg_ClosingPrice = mean(ClosingPrice)) |> 
  ggplot(aes(Avg_ClosingPrice, MonthName, fill = MonthName)) +
  geom_col() +
  facet_wrap(. ~ Year) +
  labs(
    x = "Closing Price",
    y = "Month",
    title = "Average Monthly Closing Price of Etherum"
  )
```

### Trend analysis of the change in closing price of etherum over the years
```{r}
etherum |> 
  ggplot(aes(date,ClosingPrice)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Closing Price",
    title = "Daily Closing Price of Ethereum from 2017 - 2022"
  )
```

# T-test analysis
```{r}
t.test(etherum$ClosingPrice)
```

# Time Series Analysis

### Time Series Decomposition
```{r}
#Split data into training and testing data
train_etherum <- etherum |> 
  filter(Year < 2021)
test_etherum <- etherum |> 
  filter(Year > 2020)

#convert training data to time series graphics
ts_train_etherum <- ts(train_etherum[ ,"ClosingPrice"],
                       frequency = 365,
                       start = c(2017, 11),
                       end = c(2020, 12))

#stl decompostion
fit <- stl(ts_train_etherum, 
           s.window = "periodic",
           robust = TRUE)


# Decompositon plots
autoplot(fit) +
  ggtitle("STL decompostion of Etherum Closing Prices")
```


```{r}
#ARIMA modelling
set.seed(2022)
arima_fit <- auto.arima(ts_train_etherum,
                        stepwise = F,
                        approximation = F)
print(arima_fit)
#check residuals
print(checkresiduals(arima_fit))
```

# Forecasting using the log of Closing Prices, to remove autocorrelation.
```{r}
#take log of the closing prices
transformed_etherum$ClosingPrice <- log(transformed_etherum$ClosingPrice)

#Split data into training and testing data
transformed_train_etherum <- transformed_etherum |> 
  filter(Year < 2022)
transformed_test_etherum <- transformed_etherum |> 
  filter(Year > 2021)

#convert training data to time series graphics
ts_trans_train_etherum <- ts(train_etherum[ ,"ClosingPrice"],
                       frequency = 365,
                       start = c(2017, 11),
                       end = c(2021, 12))

#stl decompostion
fit <- stl(ts_trans_train_etherum, 
           s.window = "periodic",
           robust = TRUE)


# Decompositon plots
autoplot(fit) +
  ggtitle("STL decompostion of Etherum Closing Prices")
```

```{r}
#ARIMA modelling
set.seed(2022)
arima_trans_fit <- auto.arima(ts_train_etherum,
                        stepwise = F,
                        approximation = F)
print(arima_trans_fit)
#check residuals
print(checkresiduals(arima_trans_fit))
```


```{r}
# Forecast in the next two years
arima_trans_fit |> 
  forecast() |> 
  autoplot() +
  ylab("Closing Price") +
  xlab("Year") +
  ggtitle("Forecast of Closing Price from ARIMA(2,1,1)")
```

